\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-09 paper/paper.tex
updates.

\begin{abstract}
\vspace{0.05in}
Persistent key value stores are an important component of many distributed data
serving solutions with innovations targeted at taking advantage of growing flash
speeds. Unfortunately their performance is hampered by the need to maintain and
replicate a write ahead log to guarantee availability in the face of machine and
storage failures. Cyclone is a replicated log plug-in for key value stores that
systematically addresses various sources of this bottleneck. It uses a small
amount of non-volatile memory directly addressable by the CPU - such as in the
form of NVDIMMs or Intel 3DXPoint - to remove block oriented IO devices such as
SSDs from the critical path for appending to the log. It then addresses network
overheads using an implementation of the RAFT consensus protocol that is
designed around a userspace network stack to relieve the CPU of the burden of
data copies. Finally, it provides a way to efficiently map the commutativity in
key-value store APIs to the parallelism available in commodity NICs.
Cyclone is able to replicate millions of small updates per second
using only commodity 10 gigabit ethernet adapters. As a practical application,
we use it to improve the performance (and availability) of RocksDB, a popular
persistent key value store by an order of magnitude when compared to its own
write ahead log without replication.
\end{abstract}

@@

The second key contribution of this work is in showing that the throughput
bottlenecks in the network stack can be mitigated entirely in software. Existing
work on log replication using state machine replication protocols such as
Paxos~\cite{paxos} and RAFT~\cite{raft} has not - thus far - addressed high
performance replication in the local area network of a datacenter using
commodity networks.  Research focusing on addressing wide area network latencies
by reducing the number of network hops with new protocols such as Fast
Paxos~\cite{fast-paxos} or network switch modifications such as No
Paxos~\cite{nopaxos} do not address throughput bottlenecks due to network stack
related inefficiencies in the participating nodes of the quorum -- a more urgent
problem with local area replication. On the other hand, work done to address
network related inefficiencies start with the assumption that the bottleneck to
good throughput on the network is the CPU and therefore one must either resort
to offload mechanisms such as RDMA~\cite{dare, farm, faast} or offload the
protocol entirely to an FPGA~\cite{consensus_box}. We demonstrate in this paper
that this is not the case. Rather, once persistence is provided by directly
attached non-volatile memory, thereby eliminating block IO from the persistence 
step in
replication protocols, the problem becomes equivalent to multicasting the same
log entry from the leader to all followers. We design Cyclone's network stack
around well known principles in the software packet switching community. In
particular, a careful implementation of a consensus protocol that relieves the
CPU of the responsibility of data movement permits a high performance
implementation entirely in software that addresses the order of magnitude
network related performance gap in Figure~\ref{fig:problem} using only commodity
10 Gigabit ethernet.

@@

Cyclone integrates with key value stores as both a client and server side
library. Cyclone replicates the key value store across a set of replica
servers. The client side library sends \emph{all} requests to a distinguished
leader replica.  On the server side Cyclone accepts requests from clients, calls
into the key value store, and returns a response to the client. Before
executing any PUT request, Cyclone appends it to a durable log and replicates
the request to the logs of follower replicas. The request is considered
replicated once acknowledged by a majority of replicas. Follower replicas apply
updates from the log in order.

@@

\section{Network}
\label{sec:network}
Cyclone uses the RAFT~\cite{raft} consensus protocol to replicate the log of
operations across machines. Cyclone provides guarantee to the client that any
operation responded
to is durable and its results \emph{available} as long as a majority of replicas
continue to function. Cyclone's failure model therefore is \emph{fail-recovery},
where nodes that can recover from failure simply continue to participate in
the protocol. This
requires RAFT to persist every log entry \emph{before} it sends it out for
replication. Storage overheads are therefore not independent of network
overheads, and we must deal with both storage and networking overheads
simultaneously to improve the performance of replication.

>>>>

\begin{abstract}
\vspace{0.05in}
Persistent key value stores are an important component of many distributed data
serving solutions with innovations targeted at taking advantage of growing flash
speeds. Unfortunately their performance is hampered by the need to maintain and
replicate a write ahead log to guarantee availability in the face of machine and
storage failures. Cyclone is a replicated log plug-in for key value stores that
systematically addresses various sources of this bottleneck. It uses a small
amount of non-volatile memory directly addressable by the CPU - such as in the
form of NVDIMMs or Intel 3DXPoint - to remove block oriented IO devices such as
SSDs from the critical path for appending to the log. This enables it to address network
overheads using an implementation of the RAFT consensus protocol that is
designed around a userspace network stack to relieve the CPU of the burden of
data copies. Finally, it provides a way to efficiently map the commutativity in
key-value store APIs to the parallelism available in commodity NICs.
Cyclone is able to replicate millions of small updates per second
using only commodity 10 gigabit ethernet adapters. As a practical application,
we use it to improve the performance (and availability) of RocksDB, a popular
persistent key value store by an order of magnitude when compared to its own
write ahead log without replication.
\end{abstract}

@@

The second key contribution of this work is in showing that the throughput
bottlenecks in the network stack can be mitigated entirely in software, once
block IO is removed from the critical path. Existing work on log replication
using state machine replication protocols such as Paxos~\cite{paxos} and
RAFT~\cite{raft} has not - thus far - addressed high performance replication in
the local area network of a datacenter using commodity networks.  Research
focusing on addressing wide area network latencies by reducing the number of
network hops with new protocols such as Fast Paxos~\cite{fast-paxos} or network
switch modifications such as No Paxos~\cite{nopaxos} do not address throughput
bottlenecks due to network stack related inefficiencies in the participating
nodes of the quorum -- a more urgent problem with local area replication. On the
other hand, work done to address network related inefficiencies start with the
assumption that the bottleneck to good throughput on the network is the CPU and
therefore one must either resort to network offload mechanisms such as RDMA~\cite{dare,
  farm, faast} or offload the protocol entirely to an
FPGA~\cite{consensus_box}. We demonstrate in this paper that this is not the
case. Rather, once persistence is provided by directly attached non-volatile
memory, thereby eliminating block IO from the persistence step in replication
protocols, the problem becomes akin to multicasting the same log entry
from the leader to all followers. We design Cyclone's network stack around well
known principles in the software packet switching community. In particular, a
careful implementation of a consensus protocol that relieves the CPU of the
responsibility of data movement permits a high performance implementation
entirely in software that addresses the order of magnitude network related
performance gap in Figure~\ref{fig:problem} using only commodity 10 Gigabit
ethernet.

@@

Cyclone integrates with key value stores as both a client and server side
library. Cyclone replicates the key value store across a set of replica
servers. The client side library sends \emph{all} requests to a distinguished
leader replica.  On the server side Cyclone accepts requests from clients, calls
into the key value store, and returns a response to the client. Before
executing any PUT or DELETE request, Cyclone appends it to a durable log and replicates
the request to the logs of follower replicas. The request is considered
replicated once acknowledged by a majority of replicas. Follower replicas apply
updates from the log in order.

@@

\section{Network}
\label{sec:network}
Cyclone uses the RAFT~\cite{raft} consensus protocol to replicate the log of
operations across machines. Cyclone provides guarantee to the client that any
operation responded
to is durable and its results \emph{available} as long as a majority of replicas
continue to function. Cyclone's failure model is \emph{fail-recovery},
where nodes that can recover from failure simply continue to participate in
the protocol. This
requires RAFT to persist every log entry \emph{before} it sends it out for
replication. Storage overheads are therefore not independent of network
overheads, and we must deal with both storage and networking overheads
simultaneously to improve the performance of replication.
